Unsupervised Learning for Identifying Events in Active Target Experiments

### 2020-06-17 doc/Manuscript/article.tex
Rewrote data prep + other smaller edits -MPK

One of the significant considerations for the analysis of AT-TPC data is to inject machine learning methods for track identification at the best point in the analysis pipeline.
Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for analysis.
%On the opposite end of the pipeline, 
Therefore, we will incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline.. With the benefit being that more pre-processing improves the signal-to-noise ratio, possibly improving model performance.
To explore this trade-off between model performance and pre-processing bias, we performed analysis on simulated, cleaned and raw events.

@@

To emulate the real-data case, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data. We choose this partition to be $15\%$ of each class. We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, while the entire dataset which is identified as $\boldsymbol{X}_F$. Note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.

@@

\section{Conclusions and Perspectives}\label{sec{conclusion}}
% Begin added by DB (6/17/2020)
The purpose of this study is to explore the application of unsupervised learning algorithms to event identification from an active target detector. The necessity of identifying events from raw data prior to 
% End added by DB (6/17/2020)
From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification. {\color{orange}[flip this to positive ]} It is also clear that there is significant potential in both avenues explored in this work. 
In particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
Another surprising facet of the K-means clustering is its consistent performance. As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics. This result indicates that the clusters are very clearly defined in the VGG16 latent space. As we see from the non-proton clusters if figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}  that does not necessarily imply that the physical signals are correspondingly clear.

>>>>

One of the significant considerations for the analysis of AT-TPC data is to inject machine learning methods for track identification at the best point in the analysis pipeline.
Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for analysis.
%On the opposite end of the pipeline, 
Therefore, incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline.. With the benefit being that more pre-processing improves the signal-to-noise ratio, possibly improving model performance.
To explore this trade-off between model performance and pre-processing bias, we performed analysis on simulated, cleaned and raw events.

@@

For validation, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data. We choose this partition to be $15\%$ of each class. We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, while the entire dataset which is identified as $\boldsymbol{X}_F$. Note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.

@@

\section{Conclusions and Perspectives}\label{sec{conclusion}}
% Begin added by DB (6/17/2020)
The purpose of this study is to explore the application of unsupervised learning algorithms to event identification from an active target detector. The necessity of identifying events from raw data prior to full processing is becoming a major issue in the data analysis 
% End added by DB (6/17/2020)
From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification. {\color{orange}[flip this to positive ]} It is also clear that there is significant potential in both avenues explored in this work. 
In particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
Another surprising facet of the K-means clustering is its consistent performance. As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics. This result indicates that the clusters are very clearly defined in the VGG16 latent space. As we see from the non-proton clusters if figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}  that does not necessarily imply that the physical signals are correspondingly clear.
